---
output: 
  pdf_document:
    latex_engine: xelatex
mainfont: Lato
header-includes:
- \usepackage{array}
- \usepackage{enumitem}
- \usepackage{longtable}
- \usepackage{graphicx}
- \usepackage{fontspec}
- \usepackage{fontawesome}
- \usepackage{tabularx}
- \usepackage{fullpage,graphicx,tikz}
- \usepackage{tikzpagenodes}
- \usepackage{fancyhdr}
- \usepackage{xcolor}
- \usepackage{hyperref}
---
\definecolor{downloadcolor}{RGB}{30, 135, 240}
\definecolor{mygray}{RGB}{153, 153, 153}
\usetikzlibrary{calc}

\pagestyle{fancy}
\fancyhf{}
\lfoot{\textcolor{mygray}{\textit{`r format(Sys.time(), '%B, %Y')`}}}
\cfoot{\textcolor{mygray}{\textit{Ethan S. Young $\cdot$ Curriculum Vitae}}}
\rfoot{\textcolor{mygray}{\textit{\thepage}}}
\renewcommand{\headrulewidth}{0pt}

\begin{center}
\Huge{\textbf{\textit{Ethan S. Young}}} \\
\end{center}

\vspace{2ex}

\begin{minipage}{5cm}
\begin{center}
Postdoctoral Researcher \\
Department of Pyschology, Utrecht University \\
Utrecht, The Netherlands \\
\end{center}
\end{minipage}
\hfill
\begin{minipage}{5cm}
\begin{tikzpicture}[baseline=(ola.center),inner sep=0pt]
\clip (0,0)  circle (2.25cm) node (ola) {\includegraphics[width=4.5cm]{pic1.jpeg}};
\end{tikzpicture}
\end{minipage}
\hfill
\begin{minipage}{5cm}
\faEnvelope{} \href{mailto:young.ethan.scott@gmail.com}{young.ethan.scott@gmail.com}\\
\faHome{} \href{www.ethan-young.com}{ethan-young.com} \\
\faGithub{} \href{https://github.com/ethan-young}{github}\\
\end{minipage}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
library(vitae)
library(tidyverse)
library(glue)
library(googledrive)
library(readxl)
```

```{r, include=F}
# Read cv data
cv_path <- "../../data/cv-data.xlsx"
pubs <- readxl::read_excel(cv_path, sheet = 1, col_types = c("text")) %>% 
  mutate(citation = str_replace(citation,"Young, E. S.","\\\\textbf{Young, E. S.}"))
positions <- read_excel(cv_path, sheet = 2, col_types = c("text")) %>% arrange(desc(end))
education <- read_excel(cv_path, sheet = 3, col_types = c("text")) %>% arrange(desc(end))
awards    <- read_excel(cv_path, sheet = 4, col_types = c("text")) %>% arrange(desc(end))
teaching  <- read_excel(cv_path, sheet = 5, col_types = c("text")) %>% arrange(desc(year))
reviews   <- 
  read_excel(cv_path, sheet = 6, col_types = c("text")) %>% 
  arrange(journal,desc(year)) %>% 
  group_by(journal) %>% 
  mutate(
    count = 1:n(),
    journal_group = ifelse(count == 1, journal, "")
  )
```

\vspace{1ex}

# **Education**

\vspace{1ex}
\begin{color}{mygray}\hrule\end{color}
\vspace{1ex}

```{r, results='asis'}
cat(
  "\\begin{longtable}{p{1.1in}p{5in}}",
  education %>% 
    arrange(desc(end)) %>% 
    mutate_all(~str_replace_all(.,"&","\\\\&")) %>%
    mutate(last_row = ifelse(row_number()!= max(row_number()), "\\\\& \\\\","")) %>% 
    glue_data(
      .open = "<",.close = ">",.na = "",
      "<end> & \\parbox[t]{5.0in}{<degree> \\newline <institution> \\newline <subject>} \\\\",
      "<last_row>"
    ),
  "\\end{longtable}",
  sep = "\n"
)
```

\vspace{1ex}

# **Academic Positions**

\vspace{1ex}
\begin{color}{mygray}\hrule\end{color}
\vspace{1ex}

```{r, results='asis'}
cat(
  "\\begin{longtable}{p{1.1in}p{5in}}",
  positions %>% 
    mutate_all(~str_replace_all(.,"&","\\\\&")) %>% 
    mutate(last_row = ifelse(row_number()!= max(row_number()), "\\\\& \\\\","")) %>% 
    glue_data(
      .open = "<",.close = ">",.na = "-",
      "<start>-<end> & \\parbox[t]{5.0in}{<position> \\newline <institution> \\newline advisors: <people>} \\\\",
      "<last_row>"
    ),
  "\\end{longtable}",
  sep = "\n"
)
```

\vspace{1ex}

# **Honors & Awards**

\vspace{1ex}
\begin{color}{mygray}\hrule\end{color}
\vspace{1ex}

## *Awards*

```{r, results='asis'}
cat(
  "\\begin{longtable}{p{1.1in}p{5in}}",
  awards %>% 
    filter(type == "award") %>% 
    mutate_all(~str_replace_all(.,"&","\\\\&")) %>%
    mutate(last_row = ifelse(row_number()!= max(row_number()), "\\\\& \\\\","")) %>% 
    glue_data(
      .open = "<",.close = ">",.na = "",
      "<ifelse(!is.na(start),paste0(start,'-'),start)><end> & \\parbox[t]{5.0in}{<award> \\newline <title> \\newline \\textit{<institution>}} \\\\",
      "<last_row>"
    ),
  "\\end{longtable}",
  sep = "\n"
)
```

## *Fellowships*

```{r, results='asis'}
cat(
  "\\begin{longtable}{p{1.1in}p{5in}}",
  awards %>% 
    filter(type == "fellowship") %>% 
    mutate_all(~str_replace_all(.,"&","\\\\&")) %>%
    mutate(last_row = ifelse(row_number()!= max(row_number()), "\\\\& \\\\","")) %>% 
    glue_data(
      .open = "<",.close = ">",.na = "",
      "<ifelse(!is.na(start),paste0(start,'-'),start)><end> & \\parbox[t]{5.0in}{<award> \\newline <institution>} \\\\",
      "<last_row>"
    ),
  "\\end{longtable}",
  sep = "\n"
)
```

\vspace{1ex}

# **Publications**

\vspace{1ex}
\begin{color}{mygray}\hrule\end{color}
\vspace{1ex}


## *Peer-Reviewed Papers*

```{r, results='asis'}
journals <- readxl::read_excel(cv_path, sheet = 1) %>% pull(journal) %>% unique() %>% str_subset("NA", negate = T)

walk(journals, function(x){
  pubs <<- pubs %>% 
    mutate(citation = ifelse(str_detect(citation,x), str_replace(citation,x,paste0("\\\\textit{",x,"}")),citation))
})

articles <- pubs %>% 
  filter(pub_type=="journal", in_progress == 0) %>% 
  mutate(year_arrange = ifelse(year == "in press","0",year)) %>% 
  arrange(desc(year)) %>% 
  mutate(count = 1:n()) %>% 
  group_by(year) %>%
  mutate(count_year = 1:n(),
         count_auth = ifelse(str_detect(citation,"^\\\\textbf\\{Young, E. S."), count*-1, count_year),
         border     = ifelse(count_year == max(count_auth) | n()==1,"solid","none"),
         year_group = ifelse(count_auth == min(count_auth),year,"") %>% str_replace("-"," ")) %>% 
  arrange(desc(year),count_auth) %>% 
  ungroup() %>% 
  mutate(
    border   = ifelse(count == last(count),"",border),
    download = ifelse(is.na(download),"", 
                      paste0("\\href{https://www.ethan-young.com/publications/",download,"}{\\textcolor{downloadcolor}{\\faFilePdfO{} View}}")),
  ) %>% 
  select(year_group,citation,download)

cat(
  "\\noindent \\begin{longtable}{p{1.1in}p{5in}}",
  articles %>%
    mutate(number = 1:n(),
           hline      = ifelse(lead(year_group)!="" & number<max(number),"\\hline  ",""),
           year_cell  = ifelse(!is.na(year_group), paste0("\\textbf{",year_group,"}"), " "),
           citation   = str_replace_all(citation,"&","\\\\&"),
           multi_year = ifelse(!is.na(year_cell),glue(.open = "<",.close = ">","\\multicolumn{1}{r|}{<year_cell>}"),"\\multicolumn{1}{r|}{}"),
           multi      = ifelse(lead(year_group)!=""," ", "\\multicolumn{3}{}{\\vspace{2 mm}}"),
           last_row   = ifelse(row_number()!= max(row_number()), "\\\\& \\\\[-1.5em]","")) %>% 
    glue_data(
      .open = "<",.close = ">",.na = "",
      "<year_group> & \\hangindent=0.30cm <citation> \\newline <download> \\\\ ",
      "<last_row>"
    ),
  "\\end{longtable}",
  sep = "\n"
)
```

## *Book Chapters*

```{r, results='asis'}
chapters <- pubs %>% 
  filter(pub_type=="chapter") %>% 
  arrange(desc(year)) %>% 
  mutate(count = 1:n()) %>% 
  group_by(year) %>%
  mutate(count_year = 1:n(),
         count_auth = ifelse(str_detect(citation,"^\\\\textbf\\{Young, E. S."), count*-1, count_year),
         border     = ifelse(count_year == max(count_auth) | n()==1,"solid","none"),
         year_group = ifelse(count_auth == min(count_auth),year,"") %>% str_replace("-"," ")) %>% 
  arrange(desc(year),count_auth) %>% 
  ungroup() %>% 
  mutate(
    border   = ifelse(count == last(count),"",border),
    download = ifelse(is.na(download),"", 
                      paste0("\\href{https://www.ethan-young.com/cv/",download,"}{\\textcolor{downloadcolor}{\\faFilePdfO{} View}}")),
  ) %>% 
  select(year_group,citation,download)

cat(
  "\\noindent \\begin{longtable}{p{1.1in}p{5in}}",
  chapters %>%
    mutate(number = 1:n(),
           hline      = ifelse(lead(year_group)!="" & number<max(number),"\\hline  ",""),
           year_cell  = ifelse(!is.na(year_group), paste0("\\textbf{",year_group,"}"), " "),
           citation   = str_replace_all(citation,"&","\\\\&"),
           multi_year = ifelse(!is.na(year_cell),glue(.open = "<",.close = ">","\\multicolumn{1}{r|}{<year_cell>}"),"\\multicolumn{1}{r|}{}"),
           multi      = ifelse(lead(year_group)!=""," ", "\\multicolumn{3}{}{\\vspace{2 mm}}"),
           last_row   = ifelse(row_number()!= max(row_number()), "\\\\& \\\\[-1.5em]","")) %>%
    glue_data(
      .open = "<",.close = ">",.na = "",
      "<year_group> & \\hangindent=0.30cm <citation> \\newline <download> \\\\ ",
      "<last_row>"
    ),
  "\\end{longtable}",
  sep = "\n"
)
```

## *In Progress*

```{r, results='asis'}
if(pubs %>% filter(in_progress==1) %>% nrow() > 0 ){
  in_progress <- pubs %>% 
    filter(in_progress == 1) %>% 
    mutate(count = 1:n()) %>% 
    group_by(pub_type) %>%
    mutate(count_year = 1:n(),
           count_auth = ifelse(str_detect(citation,"^\\\\textbf\\{Young, E. S."), count*-1, count_year),
           pub_group  = ifelse(count_auth == min(count_auth),pub_type,"") %>% str_replace("-"," "),
           year_group = ifelse(count_auth == min(count_auth),year,"") %>% str_replace("-"," "),
           year       = year %>% str_replace("&", "\\\\&")) %>% 
    arrange(pub_type,count_auth) %>% 
    ungroup() %>% 
    select(year_group, year,pub_group,citation)
  
  cat(
    "\\noindent \\begin{longtable}{p{1.1in}p{5in}}",
    in_progress %>%
      mutate(number = 1:n(),
             citation   = str_replace_all(citation,"&","\\\\&"),
             last_row   = ifelse(row_number()!= max(row_number()), "\\\\& \\\\[-1.5em]","")) %>%
      glue_data(
        .open = "<",.close = ">",.na = "",
        "<str_replace(year_group,'-',' ')> & \\hangindent=0.30cm <citation> \\\\ ",
        "<last_row>"
      ),
    "\\end{longtable}",
    sep = "\n"
  )
}
```

\vspace{1ex}

# **Presentations**

\vspace{1ex}
\begin{color}{mygray}\hrule\end{color}
\vspace{1ex}

## *Talks (first author only)*

```{r, results='asis'}
presentations <- pubs %>% 
  filter(pub_type=="conference", pub_sub_type == "talk") %>% 
  arrange(desc(year)) %>% 
  mutate(count = 1:n()) %>% 
  group_by(year) %>%
  mutate(count_year = 1:n(),
         count_auth = ifelse(str_detect(citation,"^\\\\textbf\\{Young, E. S."), count*-1, count_year),
         border     = ifelse(count_year == max(count_auth) | n()==1,"solid","none"),
         year_group = ifelse(month == max(month),year,"") %>% str_replace("-"," ")) %>% 
  arrange(desc(year),desc(month),count_auth) %>% 
  ungroup() %>% 
  mutate(
    border   = ifelse(count == last(count),"",border),
    download = ifelse(is.na(download),"", 
                      paste0("\\href{https://www.ethan-young.com/cv/",download,"}{\\textcolor{downloadcolor}{\\faFilePdfO{} View}}")),
  ) %>% 
  select(year_group,citation,download)

cat(
  "\\noindent \\begin{longtable}{p{1.1in}p{5in}}",
  presentations %>%
    mutate(number = 1:n(),
           hline      = ifelse(lead(year_group)!="" & number<max(number),"\\hline  ",""),
           year_cell  = ifelse(!is.na(year_group), paste0("\\textbf{",year_group,"}"), " "),
           citation   = str_replace_all(citation,"&","\\\\&"),
           multi_year = ifelse(!is.na(year_cell),glue(.open = "<",.close = ">","\\multicolumn{1}{r|}{<year_cell>}"),"\\multicolumn{1}{r|}{}"),
           multi      = ifelse(lead(year_group)!=""," ", "\\multicolumn{3}{}{\\vspace{2 mm}}"),
           last_row   = ifelse(row_number()!= max(row_number()), "\\\\& \\\\[-1.5em]","")) %>% 
    glue_data(
      .open = "<",.close = ">",.na = "",
      "<year_group> & \\hangindent=0.30cm <citation> \\newline <download> \\\\ ",
      "<last_row>"
    ),
  "\\end{longtable}",
  sep = "\n"
)

```

## *Posters (first author only)*

```{r, results='asis'}
posters <- pubs %>% 
  filter(pub_type=="conference", pub_sub_type == "poster") %>% 
  arrange(desc(year)) %>% 
  mutate(count = 1:n()) %>% 
  group_by(year) %>%
  mutate(count_year = 1:n(),
         count_auth = ifelse(str_detect(citation,"^\\\\textbf\\{Young, E. S."), count*-1, count_year),
         border     = ifelse(count_year == max(count_auth) | n()==1,"solid","none"),
         year_group = ifelse(count_auth == min(count_auth),year,"") %>% str_replace("-"," ")) %>% 
  arrange(desc(year),count_auth) %>% 
  ungroup() %>% 
  mutate(
    border   = ifelse(count == last(count),"",border),
    download = ifelse(is.na(download),"", 
                      paste0("\\href{https://www.ethan-young.com/cv/",download,"}{\\textcolor{downloadcolor}{\\faFilePdfO{} View}}")),
  ) %>% 
  select(year_group,citation,download)

cat(
  "\\noindent \\begin{longtable}{p{1.1in}p{5in}}",
  posters %>%
    mutate(number = 1:n(),
           hline      = ifelse(lead(year_group)!="" & number<max(number),"\\hline  ",""),
           year_cell  = ifelse(!is.na(year_group), paste0("\\textbf{",year_group,"}"), " "),
           citation   = str_replace_all(citation,"&","\\\\&"),
           multi_year = ifelse(!is.na(year_cell),glue(.open = "<",.close = ">","\\multicolumn{1}{r|}{<year_cell>}"),"\\multicolumn{1}{r|}{}"),
           multi      = ifelse(lead(year_group)!=""," ", "\\multicolumn{3}{}{\\vspace{2 mm}}"),
           last_row   = ifelse(row_number()!= max(row_number()), "\\\\& \\\\[-1.5em]","")) %>% 
    glue_data(
      .open = "<",.close = ">",.na = "",
      " <year_group> &\\hangindent=0.30cm <citation> \\newline <download> \\\\ ",
      "<last_row>"
    ),
  "\\end{longtable}",
  sep = "\n"
)

```

\vspace{1ex}

# **Teaching & Mentoring**

\vspace{1ex}
\begin{color}{mygray}\hrule\end{color}
\vspace{1ex}

## *Instructor*

```{r, results='asis'}
cat(
  "\\begin{longtable}{p{1.1in}p{5in}}",
  teaching %>% 
    filter(position == "Instructor") %>% 
    mutate_all(~str_replace_all(.,"&","\\\\&")) %>% 
    mutate(last_row = ifelse(row_number()!= max(row_number()), "\\\\& \\\\","")) %>% 
    glue_data(
      .open = "<",.close = ">",.na = "-",
      " <semester> <year> & <title> (<course>) \\newline <institution> \\\\",
      "<last_row>"
    ),
  "\\end{longtable}",
  sep = "\n"
)
```


## *Section Leader*

```{r, results='asis'}
cat(
  "\\begin{longtable}{p{1.1in}p{5in}}",
  teaching %>% 
    filter(position == "Section Leader") %>% 
    mutate_all(~str_replace_all(.,"&","\\\\&")) %>% 
    mutate(last_row = ifelse(row_number()!= max(row_number()), "\\\\& \\\\","")) %>% 
    glue_data(
      .open = "<",.close = ">",.na = "-",
      " <semester> <year> & <title> (<course>) \\newline <institution> \\\\",
      "<last_row>"
    ),
  "\\end{longtable}",
  sep = "\n"
)
```


## *Teaching Assistant*

```{r, results='asis'}
cat(
  "\\begin{longtable}{p{1.1in}p{5in}}",
  teaching %>% 
    filter(position == "Teaching Assistant") %>% 
    mutate_all(~str_replace_all(.,"&","\\\\&")) %>% 
    mutate(last_row = ifelse(row_number()!= max(row_number()), "\\\\& \\\\","")) %>% 
    glue_data(
      .open = "<",.close = ">",.na = "-",
      " <semester> <year> & <title> (<course>) \\newline <institution> \\\\",
      "<last_row>"
    ),
  "\\end{longtable}",
  sep = "\n"
)
```

\vspace{1ex}

# **Peer Reviews**

\vspace{1ex}
\begin{color}{mygray}\hrule\end{color}
\vspace{1ex}

```{r results='asis'}
reviews %>% 
  split(.$journal) %>% 
  iwalk(function(x, y){
    cat(
      "_**", y,"**_\n\n",
      glue_data(
        x, 
        "- {lubridate::month(as.numeric(month), label = T)}. {year}, ",
        "editor: {editor}, ",
        "{ifelse(n_revisions > 1, paste(n_revisions, 'revisions'), paste(n_revisions, 'revision'))}"
      ) %>% paste(collapse = "\n"),
      sep = "", collapse = "\n\n"
    )
  })
```

\vspace{1ex}

# **Professional Memberships**

\vspace{1ex}
\begin{color}{mygray}\hrule\end{color}
\vspace{1ex}

- Society for Personality and Social Psychology (SPSP)
- Human Behavior and Evolution Society (HBES)
